---
title: "Making simple Bioconductor containers"
author: "Laurent Gatto, Dania Machlab, Ata Badr Barzegar, João F. Carrilho, Michael Stadler"
package: SimpleBiocContainer
vignette: >
  %\VignetteIndexEntry{Building a Simple Bioconductor Container}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---


## Introduction

`SimpleBiocContainer` provides a simple, user-friendly way for users
to create R/Bioconductor containers with the convenient packages,
datasets and R scripts, as well as to share those containers with the
community through [Docker Hub](https://hub.docker.com) or the [GitHub
container
registry](https://github.blog/news-insights/product-news/introducing-github-container-registry/).


Note that you must have installed [Docker](https://www.docker.com) on
your system to be able to execute `SimpleBiocContainer` functions. You
can download and install Docker Desktop
[here](https://www.docker.com/products/docker-desktop/).

## Functions

`SimpleBiocContainer` contains the following functions:

- `makeSimpleBiocContainer()`, which creates a container directory and
  the respective files;
- `buildPushDocker()`, which builds the created container and pushes it
  to Docker Hub.

Packages to include in the container are either the currently attached
packages (default) or specified by users. The files mentioned for
`makeSimpleBiocContainer()` include datasets and scripts users may
find useful to have in the container.

## Quick start

assuming we have a character array `pacs` with the packages to store
in the container and character arrays `datasets` and `scripts` with
the paths to the data and the scripts respectively, to be included, we
can create our container and name it `"my_container"` as follows:

```{r}
container_path <- makeSimpleBiocContainer(
  package = pacs,
  container = "my_container",
  data = datasets,
  script = scripts
)
```

The code chunk above then creates our container, stores it in the
working directory, and stores the path to the container directory in
`container_path`. If `package` is not specified,
`makeSimpleBiocContainer()` will store all the packages that are
loaded in the current session.

After running `makeSimpleBiocConductor()`, we can build our container
and push it to Docker Hub as shown below:

```{r}
#| eval: false
#| include: true
buildPushDocker(
  container = container_path,
  dockerUsername = "docker_username", # replace by the intended Docker username
  ghUsername = NULL
)
```

The argument `ghUsername` (`NULL` by default) allows us to push the container to
GitHub as well, being enough to feed it with a character object that contains the
username we intend to push the container to.

## Session information

```{r}
sessionInfo()
```
